---
title: "Unveiling Key Facts about COVID-19 in Germany"
subtitle: "Introduction to R for Social Scientists"
author: "Holionko Kateryna & Myronenko Andrii"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  rmdformats::downcute
runtime: shiny_prerendered
---

\pagenumbering{arabic}

# Introduction & Research Questions

The 2019 Novel Coronavirus (COVID-19) continues to spread in countries around 
the world and we cannot see the end of this pandemic. While some countries 
remove all the restrictions such as wearing masks in public places, prohibition 
on public gatherings etc., Germany has arguably remained the only European country
where people must wear a FFP2 mask in public transport. The recent news on 
COVID-19 spread in Germany has been quite concerning, as the number of cases 
started to rise again. 
So, we decided to dive deep in the coronavirus outbreak in Germany and analyse the
dynamics of cases and deaths in the country, find out the most affected German 
States as well as the most affected age groups among Germany’s population, compare
the infection rate between males and females, determine the most popular vaccine
for German citizens and understand whether the targeted vaccination rate stated 
by the German government was achieved. 
For all these research questions to be answered we used R programming language 
and its versatile powerful packages. 

# Content

The dataset we have used was obtained from [Kaggle](https://www.kaggle.com/datasets/headsortails/covid19-tracking-germany?select=covid_de.csv). It consists of three main csv files: 

1. `covid_de.csv`: COVID-19 cases and deaths which are updated daily. The original
data are being collected by Germany's [Robert Koch Institute](https://www.rki.de/EN/Home/homepage_node.html)
and can be download through the [National Platform for Geographic Data](https://npgeo-corona-npgeo-de.hub.arcgis.com)
2. `demographics_de.csv`: General Demographic Data about Germany on the federal 
state level. This data has been downloaded from Germany's Federal Office for 
Statistics (Statistisches Bundesamt) through their Open Data platform [GENESIS](https://www-genesis.destatis.de/genesis/online/data?operation=sprachwechsel&language=en).
3. `covid_de_vaccines.csv`: COVID-19 vaccination progress for Germany as a whole. 
This dataset details daily doses, broken down cumulatively by manufacturer, 
as well as the cumulative number of people having received their first and 
full vaccination.

Let us start with the setup of the working space in R.

# Setup

```{r include=FALSE, message = FALSE, warnings = FALSE}
styler::tidyverse_style()

rm(list=ls())
knitr::opts_chunk$set(echo = FALSE)
```

We need to get rid of any possible data in our cache and load all the libraries 
which are useful for our data analysis.
We will use the `tidyverse` package to get access to a great variety of packages
such as `dplyr` and `ggplot2` which helps us to clean and visualize the data. 
The package `leaflet` helps us create an interactive clustered map for a better 
coronavirus cases representation. `RColorBrewer` allows us to pick certain color
gradients and is an extremely helpful and underrated tool for EDA. Package `shiny`
is used to create a shiny markdown in order to have a better visualization of our
final report. Finally, `plotly` helps us create interactive web-based graphs.

```{r, message = FALSE, warnings = FALSE, echo=FALSE}
library(tidyverse)
library(RColorBrewer)
library(plotly)
library(shiny)
library(ggplot2)
library(leaflet, quietly = T, warn.conflicts = F)
library(scales)
```

# Import Data

Let's import 3 csv.files: `covid_de` with all the data regarding covid-19 cases,
deaths and recoveries in Germany, `demographics_de` with the information about 
Germany's population and `covid_de_vaccines` with all the vaccination data.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
#setwd(dir = "~/Desktop/Uni/R for Social Scientists")
covid_de_data <- read.csv("covid_de.csv") %>% as_tibble()
covid_de_vaccines <- read_csv("covid_de_vaccines.csv") %>% as_tibble()
demographics_de <- read_csv("demographics_de.csv") %>% as_tibble()
```

# Data Observation & Descriptive Statistics

## Structure of data{.tabset .tabset-pills}

Let's analyse the structure of the 3 datasets.
First and foremost, we need to check the structure and extent of data by using
the following commands: `dim` and `str`

### covid_de_data

```{r, warning=FALSE, message=FALSE, echo=FALSE}
dim(covid_de_data)
str(covid_de_data)
```

`covid_de_data` dataset contains more then 2 million observation with specified 
state, country, age, gender, date, cases, death and recovered numbers. Last three
columns are in numeric format, the rest is character.

### covid_de_vaccines

```{r, warning=FALSE, message=FALSE, echo=FALSE}
dim(covid_de_vaccines)
str(covid_de_vaccines)
```

`covid_de_vaccines` dataset provides some information about date of the vaccination,
amount of first and second doses, and the amount of people being vaccinated with
Moderna/Astra Zeneca/Pfizer. All but one columns are numeric, date column is in
the Date format.

### demographics_de

```{r, warning=FALSE, message=FALSE, echo=FALSE}
dim(demographics_de)
str(demographics_de)
```

`demographics_de` shows state, gender and age grop of the person as well as the 
population. First three columns are character, whereas the last is numeric.

## {-}

We can now see, whether the number of columns and rows is what we have expected
from the original document. Further, we can already see that most of our columns
are of type numeric and character.


## Row checking{.tabset .tabset-pills}

More importantly, we need to make sure the data is indeed imported in the correct
manner, by checking some of the rows. It is known that it is usually enough to
check the first, last and some arbirtrary entries of the data:

### covid_de_data

```{r, warning=FALSE, message=FALSE, echo=FALSE}
head(covid_de_data)
tail(covid_de_data)
covid_de_data[500000:500005,]
```

### covid_de_vaccines

```{r, warning=FALSE, message=FALSE, echo=FALSE}
head(covid_de_vaccines)
tail(covid_de_vaccines)
covid_de_vaccines[200:205,]
```

### demographics_de

```{r, warning=FALSE, message=FALSE, echo=FALSE}
head(demographics_de)
tail(demographics_de)
demographics_de[100:105,]
```

## {-}

Now we can see that the data is imported in a way we expected it to be

## NA Values{.tabset .tabset-pills}

Let's now check our datasets for na values

### covid_de_data

```{r, warning=FALSE, message=FALSE, echo=FALSE}
na_rows_covid_de <- apply(is.na(covid_de_data),1,sum)/ncol(covid_de_data)
plot(na_rows_covid_de,type="h")
```

There are some missing values, but overall the data set looks very clean.

Let's now create a nice little table which helps us with an overview of the 
missingness and uniqueness of values in our columns. First, we start with calculating
the  percentage of missing values per column

```{r, warning=FALSE, message=FALSE, echo=TRUE}
missing_vals <- as.matrix(covid_de_data %>% is.na %>% colSums)
missing_vals_rel <- round(100*missing_vals/nrow(covid_de_data),2)
```

Now, we can summarise our findings in a small data summary table:

```{r, warning=FALSE, message=FALSE, echo=FALSE}
data_summary <- data.frame(missing_vals, missing_vals_rel)
colnames(data_summary) <- c("NAs", "NAs_rel")
data_summary
```

<p class="comment">
**Result**: There are a tiny 0.59% of missing data in the column "age_group" and
an insignificantnumber of missing values in the column "gender". So, our `covid_de_data`
dataset looks clean enough to explore it.
</p>

### covid_de_vaccines

```{r, warning=FALSE, message=FALSE, echo=FALSE}
na_rows_covid_de_vaccines <- apply(is.na(covid_de_vaccines),1,sum)/ncol(covid_de_vaccines)
plot(na_rows_covid_de_vaccines, type="h")
```

There are no missing values in `covid_de_vaccines` dataset so we won't explore
data in every column.

### demographics_de

```{r, warning=FALSE, message=FALSE, echo=FALSE}
na_rows_demographics <- apply(is.na(demographics_de),1,sum)/ncol(demographics_de)
plot(na_rows_demographics,type="h")
```

There are no missing values in `demographics_de` dataset so we won't explore
data in every column.

## {-}

## Summary

As a first insight it is recommended to take a look at the standard statistics of
localization and dispersion such as the mean and standard deviation.

```{r, warning=FALSE, echo=FALSE, echo=FALSE}
fluidPage(

  # Sidebar layout with a input and output definitions ----
  sidebarLayout(

    # Sidebar panel for inputs ----
    sidebarPanel(

      # Input: Selector for choosing dataset ----
      selectInput(inputId = "dataset",
                  label = "Choose a dataset:",
                  choices = c("covid_de_data", "covid_de_vaccines", "demographics_de")),
    ),

    # Main panel for displaying outputs ----
    mainPanel(

      # Output: Verbatim text for data summary ----
      verbatimTextOutput("text"),

    )
  )
)

```

```{r, context = "server", echo=FALSE}
# Return the requested dataset ----
 datasetInput <- reactive({
    switch(input$dataset,
           covid_de_data = read.csv("covid_de.csv"),
           covid_de_vaccines = read.csv("covid_de_vaccines.csv"),
           demographics_de = read.csv("demographics_de.csv"))
 #          "covid_de_data" = read.csv("~/Desktop/Uni/R for Social Scientists/covid_de.csv"), #write your working directory here
#           "covid_de_vaccines" = read.csv("~/Desktop/Uni/R for Social Scientists/covid_de_vaccines.csv"),
#           "demographics_de" = read.csv("~/Desktop/Uni/R for Social Scientists/demographics_de.csv"))
  })

  # Generate a summary of the dataset ----
  output$text <- renderPrint({
    dataset <- datasetInput()
    summary(dataset)
  })

```

This brief summary of course is nothing more than a quick peek at the data. So, 
finally we are ready to do exploratory data analysis in relation to our research
questions.

# Exploratory Data Analysis (EDA) 

## COVID-19 case and death progression by German State{.tabset .tabset-pills}

To begin with, we would like to see the dynamics of coronavirus cases and
deaths in each German State over time. So let's create a line graphs with COVID-19
cases (we will draw this graph from November 2020, not the date of the very first
recorded COVID-19 case, since the dynamics of case progression is better
seen this way) and deaths over time.

### Case progression

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_de_data %>%
  arrange(date) %>%
  group_by(date, state) %>%
  summarise(cases = sum(cases),
            deaths = sum(deaths),
            recovered = sum(recovered)) %>%
  ungroup() %>%
  group_by(state) %>%
  summarise(date = date,
            cases = cases,
            cumulative_cases = cumsum(cases),
            deaths = deaths,
            cumulative_deaths = cumsum(deaths),
            recovered = recovered,
            cumulative_recoveries = cumsum(recovered)) %>%
  ungroup() %>%
  filter(date > "2020-11-01") %>%
  plot_ly(x = ~date,
          y = ~cumulative_cases,
          type = "scatter",
          mode = "lines",
          color = ~state) %>%
  layout(yaxis = list(title = "Cases"),
         xaxis = list(title = "Date"))
```

### Death progression

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_de_data %>%
  arrange(date) %>%
  group_by(date, state) %>%
  summarise(cases = sum(cases),
            deaths = sum(deaths),
            recovered = sum(recovered)) %>%
  ungroup() %>%
  group_by(state) %>%
  summarise(date = date,
            cases = cases,
            cumulative_cases = cumsum(cases),
            deaths = deaths,
            cumulative_deaths = cumsum(deaths),
            recovered = recovered,
            cumulative_recoveries = cumsum(recovered)) %>%
  ungroup() %>%
  plot_ly(x = ~date,
          y = ~cumulative_deaths,
          type = "scatter",
          mode = "lines",
          color = ~state) %>%
  layout(yaxis = list(title = "Deaths"),
         xaxis = list(title = "Date"))
```

## {-}

From the "Cases progression by German State" chart we can conclude that three 
German States, namely Nordrhein-Westfallen, Bayern and Baden-Wuerttemberg have far
higher coronavirus cases than the rest of Germany. Regarding deaths progression 
Bayern and Nordrhein-Westfallen have the worst situation, however, the number 
of deaths in Sachsen, which does not have as many infections as Bayern, 
Nordrhein-Westfallen and Baden-Wuerttemberg, is almost the same as in Baden-Wuerttemberg.

## German States with the most coronavirus cases and deaths{.tabset .tabset-pills}

Let's breakdown German States by the largest total number of coronavirus cases 
and deaths. 

### States with the greatest number of cases

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_de_data %>%
  arrange(date) %>%
  group_by(date, state) %>%
  summarise(cases = sum(cases),
            deaths = sum(deaths),
            recovered = sum(recovered)) %>%
  ungroup() %>%
  group_by(state) %>%
  summarise(date = date,
            cases = cases,
            cumulative_cases = cumsum(cases),
            deaths = deaths,
            cumulative_deaths = cumsum(deaths),
            recovered = recovered,
            cumulative_recoveries = cumsum(recovered)) %>%
  ungroup() %>%
  filter(date == date[length(date)]) %>%
  arrange(desc(cumulative_cases)) %>%
  top_n(7) %>%
  plot_ly(x = ~cumulative_cases,
          y = ~reorder(state, cumulative_cases),
          type = 'bar',
          orientation = 'h',
          marker = list(color = c("orangered", "orangered", "orangered", 
                                  "lightskyblue", "lightskyblue", "lightskyblue", 
                                  "lightskyblue"))) %>%
  layout(yaxis = list(title = ""),
         xaxis = list(title = "Cases"))
```

### States with the greatest number of deaths

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_de_data %>%
  arrange(date) %>%
  group_by(date, state) %>%
  summarise(cases = sum(cases),
            deaths = sum(deaths),
            recovered = sum(recovered)) %>%
  ungroup() %>%
  group_by(state) %>%
  summarise(date = date,
            cases = cases,
            cumulative_cases = cumsum(cases),
            deaths = deaths,
            cumulative_deaths = cumsum(deaths),
            recovered = recovered,
            cumulative_recoveries = cumsum(recovered)) %>%
  ungroup() %>%
  filter(date == date[length(date)]) %>%
  arrange(desc(cumulative_deaths)) %>%
  top_n(7) %>%
  plot_ly(x = ~cumulative_deaths,
          y = ~reorder(state, cumulative_deaths),
          type = 'bar',
          orientation = 'h',
          marker = list(color = c("orangered", "orangered", "orangered", 
                                  "orangered", "lightskyblue", "lightskyblue", 
                                  "lightskyblue"))) %>%
  layout(yaxis = list(title = ""),
         xaxis = list(title = "Deaths"))
```

## {-}

Now we can clearly see that Nordrhein-Westfallen, Bayern and Baden-Wuerttemberg
have the most number of infections and consequently the largest number of covid-
related deaths. Although Sachsen registered more than 2 times less coronavirus cases
than Baden-Wuerttemberg, its death toll turned out to be almost the same.

Let's now create a table with the total number of cases and deaths by age groups.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_data_by_age <- covid_de_data %>%
  na.omit() %>%
  arrange(date) %>%
  group_by(date, state, age_group) %>%
  summarise(cases = sum(cases),
            deaths = sum(deaths),
            recovered = sum(recovered)) %>%
  ungroup() %>%
  group_by(age_group) %>%
  summarise(Cases = sum(cases),
            Deaths = sum(deaths)) %>%
  ungroup()

covid_data_by_age
```

Now our data is suitable for doing meaningful charts to analyse cases and deaths
in different age groups.

## Breakdown of coronavirus cases and deaths by age group{.tabset .tabset-pills}

### Total coronavirus cases by age group

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_data_by_age %>%
  plot_ly(x = ~age_group,
          y = ~Cases,
          type = "bar") %>%
  layout(yaxis = list(title = "Cases"),
         xaxis = list(title = "Age Group"))
```

### Total coronavirus deaths by age group

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_data_by_age %>%
  plot_ly(x = ~age_group,
          y = ~Deaths,
          type = "bar",
          marker = list(color = c("lightskyblue", "lightskyblue", "lightskyblue",
                                  "lightskyblue", "darkorange", "orangered"))) %>%
  layout(yaxis = list(title = "Deaths"),
         xaxis = list(title = "Age Group"))
```

## {-}

It is evident from the charts that the most number of coronavirus cases is recorded
in the age group from 35 to 59, while the greatest number of deaths occur in 
80-99 age group, despite this group has the second lowest number of cases.

## Death Rate by age group

Let's calculate the death rate for each age group and depict it.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_data_by_age %>%
  mutate(Death_Rate = round(Deaths/Cases * 100, 2)) %>%
  plot_ly(x = ~age_group,
          y = ~Death_Rate,
          type = "scatter",
          mode = "lines+markers") %>%
  layout(xaxis = list(title = "Age Group"),
         yaxis = list(title = "Death Rate",
                      ticksuffix = "%"))
```

From this graph we may make a conclusion that the older a person is, the more
likelihood there is that the person dies from coronavirus. Almost every tenth person
aged 80-99 dies from COVID-19 as the diagram suggests.

For further analysis we would like to create a table with the total number of 
cases and deaths among males and females by different age groups and German States.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_data_by_age_state <- covid_de_data %>%
  na.omit() %>%
  arrange(date) %>%
  group_by(date, state, age_group, gender) %>%
  summarise(cases = sum(cases),
            deaths = sum(deaths),
            recovered = sum(recovered)) %>%
  ungroup() %>%
  group_by(age_group, state, gender) %>%
  summarise(Cases = sum(cases),
            Deaths = sum(deaths)) %>%
  ungroup()

covid_data_by_age_state
```

For a better visuals let's change the data in the column "gender" - specify that
"F" is "female" and "M" is "male".

```{r, warning=FALSE, message=FALSE, echo=TRUE}
covid_data_by_age_state <- covid_data_by_age_state %>%
  mutate(gender = recode(gender, "F" = "female", "M" = "male"))

covid_de_data <- covid_de_data %>%
  mutate(gender = recode(gender, "F" = "female", "M" = "male"))
```

The data in our tables now looks consistent and analyzable. Let's now look at the
relative number of infected people. For this purpose let's join the column "population"
from `demographics_de` datasetand calculate the percent of infected people by 
each age_group.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_data_by_age_state <- left_join(covid_data_by_age_state,
                                     demographics_de,
                                     by = c("state", "age_group", "gender"))
covid_data_by_age_state
```

Finally, we can create a vertical bar chart for the proportion of infections
in different age group to a total population of this age group.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_data_by_age_state %>% 
  group_by(age_group) %>%
  summarise(Cases = sum(Cases),
            Deaths = sum(Deaths),
            Population = sum(population)) %>%
  ungroup() %>%
  mutate(infection_rate = round(Cases/Population * 100 , 2)) %>%
  plot_ly(x = ~age_group,
          y = ~infection_rate,
          type = "bar") %>%
  layout(yaxis = list(title = "Infection Rate",
                      dtick = 10, 
                      tick0 = 0, 
                      ticksuffix = "%",
                      range = c(0, 55)),
         xaxis = list(title = "Age Group"))
```

From the bar chart it is evident that more than half of children aged 5-14 got 
infected with COVID-19. This is followed by people aged 15-34 and 35-59 
respectively. So, a lot of children and young adults caught COVID-19, although
the death rate in these age groups is substantially low.

## Coronavirus cases and deaths distribution by gender{.tabset .tabset-pills}

Now we would like to compare number of infections and deaths among males and 
females. For this purpose we will create pie charts.

### Covid-19 cases distribution by gender

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_de_data %>%
  na.omit() %>%
  arrange(date) %>%
  group_by(date, state, age_group, gender) %>%
  summarise(cases = sum(cases),
            deaths = sum(deaths),
            recovered = sum(recovered)) %>%
  ungroup() %>% 
  group_by(gender) %>%
  summarise(Cases = sum(cases),
            Deaths = sum(deaths)) %>%
  ungroup() %>%
  plot_ly(labels = ~gender,
          values = ~Cases,
          type = "pie",
          textposition = "inside",
          textinfo = "label+percent",
          showlegend = FALSE,
          marker = list(colors = c("#FF9933",
                                   "#009900"))) %>%
  layout(title = 'Cases Distribution by gender',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
```

### Covid-19 deaths distribution by gender

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_de_data %>%
  na.omit() %>%
  arrange(date) %>%
  group_by(date, state, age_group, gender) %>%
  summarise(cases = sum(cases),
            deaths = sum(deaths),
            recovered = sum(recovered)) %>%
  ungroup() %>% 
  group_by(gender) %>%
  summarise(Cases = sum(cases),
            Deaths = sum(deaths)) %>%
  ungroup() %>%
  plot_ly(labels = ~gender,
          values = ~Deaths,
          type = "pie",
          textposition = "inside",
          textinfo = "label+percent",
          showlegend = FALSE,
          marker = list(colors = c("#FF9933",
                                   "#009900"))) %>%
  layout(title = 'Deaths Distribution by gender',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
```

## {-}

From these pie charts we can clearly see that although the number of women, 
who caught coronavirus, is a few percents larger, the proportion of deaths 
turned out to be 6.6% higher for males

## Comparison of infection rates among male and female of different age groups

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_data_by_age_state %>% 
  group_by(age_group, gender) %>%
  summarise(Cases = sum(Cases),
            Deaths = sum(Deaths),
            Population = sum(population)) %>%
  ungroup() %>%
  mutate(infection_rate = Cases/Population) %>%
  ggplot(aes(fill=gender, y=infection_rate, x=age_group)) +
  geom_bar(position = "dodge", stat = "identity") +
  scale_y_continuous(breaks = seq(0, 0.6, by = 0.1), labels = scales::percent) +
  scale_fill_brewer(palette = "Accent") +
  labs(title = "Infection Rate Dynamics of men and women by age groups",
       x = "Age Group",
       y = "Infection Rate") +
  theme_light() +
  geom_text(aes(label = percent(infection_rate)), vjust = 2.5, 
            position = position_dodge(.9), colour = "black", size = 2.5)
```

It is evident from this grouped bar chart that in the age group younger than 15 years old
there are more males infected. From 15 to 59 years old the proportion of females 
who caught COVID-19 is few percents higher than that of men. In the most vulnerable
age group 80-99 the proportions are distributed almost equally.

## Coronavirus cases distribution map view

Let's show the comparison of the total coronavirus cases in each German State 
on the map. For this purpose we need to add coordinates of each State.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
latitude <- c(48.99999229, 50.26660748, 52.52181866, 51.7704175, 53.55043805, 53.55002464,
              51.30000694, 53.63330408, 53.1299986, 52.02998822, 49.98247246, 49.25039044,
              50.82998395, 52.13042137, 54.78374778, 50.97005292)

longtitude <- c(8.399993448, 10.96660681, 13.40154862, 14.32996741, 8.579982461,
                9.999999144, 9.500029662, 11.41669861, 8.220004434, 8.530011351,
                8.273219156, 6.970003213, 12.91997595, 11.62000362, 9.433315388,
                11.02996212)
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_data_map <- covid_data_by_age_state %>%
  group_by(state) %>%
  summarise(Cases = sum(Cases),
            Deaths = sum(Deaths),
            population = sum(population)) %>%
  ungroup()
```

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_data_map$lat <- latitude
covid_data_map$long <- longtitude
covid_data_map
```

And ultimately we can create a map with clusters, each corresponding to a certain
German state and reflecting the total number of coronavirus cases recorded in this
state.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
library(leaflet)
map <- leaflet() %>%
  addProviderTiles(providers$Esri.WorldStreetMap) %>%
  
  setView(lng = 10.4515, lat = 51.7657, zoom = 6) %>%
  
  addCircleMarkers(data = covid_data_map,
                   covid_data_map$long,
                   covid_data_map$lat,
                   weight = 0.5,
                   color = "green",
                   radius = covid_data_map$Cases/125000,
                   fillOpacity = 0.7,
                   stroke = T,
                   label = ~paste0(state, ": ", 
                                   as.character.numeric_version(Cases)), 
                   group = "Points")

map
```

## Vaccine Analysis

Here we are going to explore our third dataset, `covid_de_vaccines`, to analyse
the speed of vaccination by each of three vaccines. Unfortunately, the data given
is limited to the cumulative number of vaccines and date of first and second doses,
so we were not able to join the dataset with the previous ones. However, it is
still interesting to know which vaccine was known to be the most popular one and
how the trend has changed.  

```{r, warning=FALSE, message=FALSE, echo=FALSE}
covid_de_vaccines %>%
  plot_ly(x = ~date,
          y = ~pfizer_cumul,
          name = "Pfizer",
          type = "scatter",
          mode = "lines") %>%
  add_trace(y = ~astrazeneca_cumul, name = "Astra Zeneca") %>%
  add_trace(y = ~moderna_cumul, name = "Moderna") %>%
  layout(title = "Vaccinations progression over time",
         xaxis = list(title = "Date"),
         yaxis = list(title = "Number of jabs administered"))
```

From the line graph we may clearly see the popularity of Pfizer vaccine, which
more than 130 million people received by June 2022. Only around 30 million people
were vaccinated with Moderna, which is followed by nearly 13 million getting 
Astra Zeneca vaccine. Notably, we may observe a speed curve of Pfizer vaccination
starting from February 2021.
From our perspective, such a high popularity of Pfizer may be explained by the
decision of German government make a Pfizer vaccine a default option and therefore
German citizens simply did not have a choice to choose the vaccine producer. 
Moreover, after some time since launching Astra Zeneca was known to have many 
side-effects and therefore the usage of this vaccine type was decreased dramatically.

## Targeted vaccination rate

Now we want to check if the target vaccination rate of 90% stated by the German
government was achieved. At first let's calculate total German population.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
germany_population <- sum(demographics_de$population)
germany_population
```

We can now create a bar gauge for the current vaccination rate in Germany,
namely the relative number of people who got the first jab and the percent of 
citizens who received a full vaccination - 2 doses of vaccines.

```{r, warning=FALSE, message=FALSE, echo=FALSE, fig.align="center"}
covid_de_vaccines %>%
  plot_ly(type = "indicator",
            mode = "number+gauge+delta",
            value = ~persons_full_cumul[531]/germany_population * 100,
            domain = list(x = c(0.25, 1), y= c(0.4, 0.6)),
            title = list(text = "<b>Full vaccination</b>"),
            delta = list(reference = 90),
            gauge = list(shape = "bullet",
                         axis = list(range = list(NULL, 100)),
            threshold = list(line = list(color = "black", width = 3),
                             thickness = 1,
                             value = 90),
            steps = list(list(range = c(0, 50), color = "lightgreen"),
                         list(range = c(50, 85), color = "green")),
            bar = list(color = "black"))) %>%
  add_trace(type = "indicator",
          mode = "number+gauge+delta",
          value = ~persons_first_cumul[531]/germany_population * 100,
          domain = list(x = c(0.25, 1), y= c(0.7, 0.9)),
          title = list(text = "<b>First Dose</b>"),
          delta = list(reference = 90),
          gauge = list(shape = "bullet",
                       axis = list(range = list(NULL, 300)),
          threshold = list(line = list(color = "black", width = 3),
                           thickness = 1,
                           value = 90),
          steps = list(list(range = c(0, 50), color = "lightgreen"),
                       list(range = c(50, 85), color = "green")),
          bar = list(color = "black")))

```

From these graphs we may conclude that 77.6% of the German population received one
dose of the vaccine and 76% are fully vaccinated. However, the target of 90% 
set by the German government was not yet achieved. Thus, at the moment almost 
every fourth person has not received a jab against COVID-19, so German population
has not developed a herd immunity yet and therefore the number of coronavirus 
cases started to rise again this summer.

# Conclusion

After a detailed exploratory data analysis of our dataset we have come to the following
conclusions:

1. The most COVID-19 affected German States turned out to be the most populated
districts - Nordrhein-Westfallen, Bayern and Baden-Wuerttemberg. Although Sachsen
has a much lower population and number of recorded cases than Baden-Wuerttemberg
its death toll appeared to be almost the same as in this State, most likely because
of the ageing population.
2. The most infected age group is the largest one - people aged from 35 to 59 years
old, while the highest mortality rate is observed among the elderly people - 
80-99. So, the older a person is the higher the likelihood that this person dies
from coronavirus.
3. The proportion of females that caught COVID-19 is few percents higher, since
the number of women living in Germany is higher than that of men; however, roughly
7% more males died from coronavirus.
4. Among 3 different vaccine producers, the number of people that received Pfizer
jab is significantly higher than that of Moderna and Astra Zeneca and makes up
more than 130 million doses administered by June 10, 2022. It is also notable, that
Astra Zeneca vaccines were discontinued in July 2021 due to discovered rare
blood clots.
5. The targeted vaccination rate of 90% stated by the German government was not 
yet achieved. Current figures stand at 76% for fully vaccinated German population,
which means that almost every fourth person in Germany has still not received
a jab. That is why the number of coronavirus cases started to surge again this
summer.